---
title: "ActivityMonitoring"
author: "Richard Ellison"
date: "Wednesday, June 10, 2015"
output: html_document
---

#Reproducible Research - Peer Assessment 1#

##Load the data and transform for later analysis##
```{r setup, cach=FALSE, include=FALSE}
library(knitr)
opts_chunk$set(fig.path='figure/')
```


```{r, echo=TRUE}
#DataFile <- ("./Coursera/ActivityMonitoring/activity.csv")
DataFile <- "activity.csv"
ActivityData <- read.csv(DataFile)

# format string to date format
ActivityData$date <- as.POSIXct(ActivityData$date, format="%Y-%m-%d")

# Build dataframe for weekday work
ActivityData <- data.frame(date=ActivityData$date,
                        steps=ActivityData$steps,
                        interval=ActivityData$interval,
                        weekday=weekdays(ActivityData$date))

# Compute the daytype from the weekday from the date
ActivityFull <- cbind(ActivityData, daytype=ifelse(ActivityData$weekday=="Saturday"|
                        ActivityData$weekday=="Sunday","weekend","weekday"))

# clear out the original
#rm(ActivityData)

```

**What is the mean total of steps taken per day?  First show histogram of all steps.**

```{r, echo=TRUE}
StepsPerDay <- aggregate(steps~date, ActivityFull, FUN=sum)
#histogram for total number of steps taken each day
barplot(StepsPerDay$steps, names.arg=StepsPerDay$date, xlab="date", ylab="steps")
```

**Then report on mean and median steps per day**

```{r, echo=TRUE}
mean(StepsPerDay$steps)
median(StepsPerDay$steps)
```

**What is the average daily activity pattern?**

```{r, echo=TRUE}
#Make a time series plot of the 5-minute interval and avg num of steps,
#    averaged across all days
AverageSteps <- aggregate(steps~interval, ActivityFull, FUN=mean)
plot(AverageSteps, type="l")
```

**Which 5-minute interval, on average, contains the maximum number of steps?**

```{r, echo=TRUE}
AverageSteps$interval[which.max(AverageSteps$steps)]
```

**Find number of rows with missing values (NA)**
```{r, echo=TRUE}
sum(is.na(ActivityFull))

```

**Impute missing values from 5-minute intervals**
```{r, echo=TRUE}
ActivityFull <- merge(ActivityFull, AverageSteps, by="interval")
NARecords <- is.na(ActivityFull$steps.x)
ActivityFull$steps[NARecords] <- ActivityFull$steps.y[NARecords]
ActivityFull <- ActivityFull[, c(1,2,4,5,6)]
names(ActivityFull)[names(ActivityFull)=="steps.y"] <- "steps"
#check that no nulls remain
sum(is.na(ActivityFull))

```

**Make a histogram of the total number of steps taken each day**
```{r, echo=TRUE}
StepsPerDay2 <- aggregate(steps~date, ActivityFull, FUN=sum)
#barplot(StepsPerDay2$steps, names.arg=StepsPerDay2$date, xlab="date",
barplot(StepsPerDay$steps, names.arg=StepsPerDay$date, xlab="date", 
        ylab="steps", ylim=c(0,20000))
```

**Show mean and median total number of steps per day**
```{r, echo=TRUE}
mean(StepsPerDay2$steps)
median(StepsPerDay2$steps)
```

*Missing data has a low impact between missing data and estimates*

**Are there differences in activity patterns between weekdays and weekends?**
```{r, echo=TRUE}
par(mfrow=c(2,1))
for (type in c("weekend","weekday")) {
        StepsType <- aggregate(steps~interval, data=ActivityFull, 
                                subset=ActivityFull$daytype==type,
                                FUN=mean)
        plot(StepsType, type="l", main=type)
}
```

echo=FALSE
#```{r, echo=FALSE}
#par(mfrow=c(2,1))
#StepsByDayFinal <- aggregate(steps ~ interval + daytype, ActivityFull, mean)
#with(StepsByDayFinal, plot(steps~interval, type="n", main="Weekday vs. Weekend"))
#with(StepsByDayFinal[StepsByDayFinal$daytype=="weekday",], lines(steps~interval, type="l", col="red"))
#with(StepsByDayFinal[StepsByDayFinal$daytype=="weekend",], lines(steps~interval, type="l", col="blue"))
#```
#``{r, echo=FALSE}
#par(mfrow=c(2,1))
#StepsByDayFinal <- aggregate(steps ~ interval + daytype, ActivityFull, mean)
#StepsByDayWeekday <- subset(StepsByDayFinal, daytype=c("weekday"))
#StepsByDayWeekend <- subset(StepsByDayFinal, daytype=c("weekend"))
#WeekdaySteps <- aggregate(steps~interval, StepsByDayWeekday, FUN=mean)
#plot(WeekdaySteps, type="l")
#WeekendSteps <- aggregate(steps~interval, StepsByDayWeekend, FUN=mean)
#plot(WeekendSteps, type="l")
#```

